Running simulations by Alisa

### 2015-10-23 sims/run-sim-Alisa.Rmd
peter fiddling

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
source("sim-fns.R")
```

@@

Now to define selection. Here we selection against heterozygotes to be strong (0.5).
```{r set_selection}
	S  = 0.5
	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
```
`qtl` is a named list, one element per chromosome,
and each element of the list is a data frame
whose columns give the name of the locus,
the selection coefficient,
and the position in Morgans.

@@

This is a two-step process. First we simulate genotypes:
```{r run_sim_step1}
	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
	#This can take a while, 
```
The output is: a list with hierarchy: [[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]
	where:
		"inds" is a list of length ninds, such that [[ind]][[chr]][[X1/X2]][[data.frame where col1 = coordinates of rec., col2 = ancestor ID]]
		"sp.inds" is a character vector of length ninds, giving the ancestry given an ancestor ID
		"deme" is an integer vector of length ninds, giving the deme number of an ancestor, given their ancestor ID
		"pars" is an integer that gives the number of generatsion of the simulation
		"QTL" returns the QTL vector used in the simulation

The second step is to parse the simulated genotypes:
```{r run_sim_step2}
	sims.sums = lapply(sims.s0.1,spBreaks)
```
The output here is a list: [[zone_age]][[c("ind.ancestry","ancest.prop")]]
		"ind.ancestry" is a list of length ninds: [[ind][[chr]]][[X1/X2]][[data.frame where col1 = (chunk)starts, col2 = (chunk)stops, col3 = sp2(T/F)]]
		"ancest.prop" is a ninds*2 matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.

This resulting files can get quite large, but we've saved them here
```{r save_file,cache=TRUE}
	save(sims.s0.1,file="sims_sums_s0.5_sigma1_50by500_tau50.Robj")
```

@@

Now, we get frequencies at each of the sites in the above vector, 
first by getting genotypes at all loci, then using those to calculate an allele frequency for each deme:
```{r get_frequencies}
	genotypes = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
	freqs = lapply(genotypes,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
	xx <- (1:ndemes)-0.5-ndemes/2
		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")

```
(`genotypes` is a list of nloci*ninds dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
Each line in the plot represents a locus a given distance from the selected locus.

@@

```{r peters_predictions}
	zeta <- function(x,r){
	2*pnorm(abs(x)) - 1 + 2 * exp(r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE))	
	}

	pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
	(1/2)*(1-sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau))	
	}

@@

		#function to increase transparency of colors
		add.alpha <- function(col, alpha=1){
		if(missing(col))
		stop("Please provide a vector of colours.")
		apply(sapply(col, col2rgb)/255, 2, 
		function(x) 
		rgb(x[1], x[2], x[3], alpha=alpha)) 
		}

	transparent_rainbow = add.alpha(rainbow(ndemes),0.75)

		D = 1; hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")

		for(D in seq(1,ndemes,5)){
		hist(tracts[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
		}
```
To get LD in these populations, we define `get.genotype` to get the genotypes at two loci in individuals, and `get.LD` to calculate LD in a deme:
```{r LD_functions}
 	get.genotype = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS1=0.5,POS2=0.5){
	site1 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS1 & X$stops>POS1),3]})	
	site2 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS2 & X$stops>POS2),3]})
	return(rbind(site1,site2))
	}

@@

```
LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:ndemes,get.LD,POS1=0.5,POS2=Z,CHR=1,IND_DATA=sims.sums[[1]]$ind.ancest)}))
	matplot(xx,LD_matrix,type="l",col=rainbow(20),xlim=c(-10,10),lty=1)

>>>>

```{r doc_setup, include=FALSE}
fig.dim <- 5
knitr::opts_chunk$set(fig.height=fig.dim,fig.width=2*fig.dim,fig.align='center')
```

@@

Now to define selection. 
```{r set_selection}
	S  = 0.5
	qtl=list(chr1=data.frame(traits=c("underdominant"), s = S,pos=c(0.5)))
```
Here we selection against heterozygotes to be strong (`r S`).
`qtl` is a named list, one element per chromosome,
and each element of the list is a data frame
whose columns give the name of the locus,
the selection coefficient,
and the position in Morgans.

@@

This is a two-step process. First we simulate genotypes:
```{r run_sim_step1,cache=TRUE,depends="set_selection"}
	sims.s0.1 = lapply(zone_age,sim.zone,n.ind=ninds,n.deme=ndemes,sigma=SIGMA)
	#This can take a while, 
```
The output is: a list with hierarchy: `[[zone_age]][[c("inds","sp.inds","deme","pars","QTL")]]` where:

	- `inds` is a list of length `ninds`, such that each `[[ind]][[chr]][[X1/X2]]` is a data.frame with columns
        * `pos` : coordinates of rec.,  (*what does this mean?*)
        * `ancest` : ancestor ID
	- `sp.ids` is a character vector of length `ninds`, giving the ancestry given an ancestor ID
	- `deme` is an integer vector of length `ninds`, giving the deme number of an ancestor, given their ancestor ID
	- `pars` is an integer that gives the number of generations of the simulation
	- `QTL` is the QTL vector used in the simulation

The second step is to parse the simulated genotypes:
```{r run_sim_step2,cache=TRUE,depends="run_sim_step1"}
	sims.sums = lapply(sims.s0.1,spBreaks)
```
The output here is a list: `[[zone_age]][[c("ind.ancestry","ancest.prop")]]`

    * `ind.ancestry` is a list of length ninds: each `[[ind][[chr]]][[X1/X2]]` is a data.frame with columns
        * `starts` = (chunk) starts, 
        * `stops` = (chunk) stops, 
        * `sp2` = sp2(T/F) (*what is this?*)
    * `ancest.prop` is a $2 \times$`ninds` matrix giving the ancestry proportion of each individual on each of their two copies of the chromosome.

This resulting files can get quite large, but we've saved them here
```{r save_file,cache=TRUE,depends="run_sim_step1"}
	save(sims.s0.1,file="sims_sums_s0.5_sigma1_50by500_tau50.Robj")
```

@@

Now, we get frequencies at each of the sites in the above vector, 
first by getting genotypes at all loci, then using those to calculate an allele frequency for each deme:
```{r get_frequencies}
	genotypes = lapply(sims.sums,function(Z){data.frame(do.call(cbind,lapply(Z$ind.ancest, geno.ind, loci)))})
	freqs = lapply(genotypes,function(X){apply(X,1,function(Z){tapply(Z,cut(1:ncol(X),breaks=seq(0,ncol(X),deme_size)),mean)/2})})
	xx <- (1:ndemes)-0.5-ndemes/2
		matplot(xx,freqs[[1]],col=rainbow(20),lty=1,type="l", main="50 generations",ylab="freq",xlab="deme")
		legend('bottomright',col=rainbow(20),lty=1,legend=loci[[1]],cex=0.75)
		matpoints(-xx,pp,col=rainbow(20),lty=3,type="l")
```
(`genotypes` is a list of `nloci`$\times$`ninds` dataframes where each entry is the number of ancestry B alleles in an individual at the given locus.)
Each line in the plot represents a locus a given distance from the selected locus.

@@

```{r peters_predictions}
    zeta <- function(x,r){
        2*pnorm(abs(x)) - 1 + 2 * exp(r*abs(x) + r^2/2 + pnorm(abs(x)+r,lower.tail=FALSE,log.p=TRUE))	
    }
    pcline <- function(x,r,tau=zone_age,sigma=SIGMA){
        (1/2)*(1-sign(x)*zeta(x=x/sqrt(tau*sigma^2),r=r*tau))	
    }

@@

	transparent_rainbow = adjustcolor(rainbow(ndemes),0.75)

    D = 1 
    hist(tracts[[D]], border="black",breaks=seq(0,1,0.05),ylim=c(0,100),xlab="length",main="Distribution of tracts")
    for (D in seq(1,ndemes,5)) {
        hist(tracts[[D]], col=transparent_rainbow[D],border=NA,breaks=seq(0,1,0.05),ylim=c(0,50),add=T)	
    }
```
To get LD in these populations, we define `get.genotype` to get the genotypes at two loci in individuals, and `get.LD` to calculate LD in a deme:
```{r LD_functions}
 	get.genotype = function(IND_DATA=sims.sums[[1]]$ind.ancest[[1]],CHR=1,POS1=0.5,POS2=0.5){
	site1 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS1 & X$stops>POS1),3]})	
	site2 = sapply(IND_DATA[[CHR]],function(X){X[which(X$starts<POS2 & X$stops>POS2),3]})
	return(rbind(site1,site2))
	}

@@

```{r ld_matrix}
LD_matrix = do.call(cbind, lapply(seq(0.51,0.99,0.05),function(Z){sapply(1:ndemes,get.LD,POS1=0.5,POS2=Z,CHR=1,IND_DATA=sims.sums[[1]]$ind.ancest)}))
	matplot(xx,LD_matrix,type="l",col=rainbow(20),xlim=c(-10,10),lty=1)
